A Large and Variable Leading Tail of Helium in a Sub-Saturn Undergoing Runaway Inflation

### 2023-07-06 paper/paper1/ms.tex
define rho XUV

\begin{equation}
    \dot{M} = \frac{3}{4}\frac{\eta F_\mathrm{XUV}}{G K_t \rho_\mathrm{XUV}}\label{thorn23Mdot}.
\end{equation}
 where $K_t$ is the tidal gravity term, and $\eta \sim 0.4$ comes from \citet{2022A&A...663A.122C}.  Assuming $\log{L_\mathrm{XUV}/L_\mathrm{bol}}=-4.2$, we would expect \hatpb~ to exhibit $\dot{M}\sim10^2$ M$_\oplus$/Gyr. At a current $M\sim95\;$M$_\oplus$ and an accelerating mass loss rate, its lifetime would measure in the $\sim$100 Myr range.  The \citet{2023ApJ...945L..36T} simulations focused on 0.75-1.25 $M_\odot$ host stars; the higher mass $\sim$1.6 $M_\odot$ \hatp would deliver greater tidal gravity for a given semi-major axis, and possibly lower $\log{L_\mathrm{XUV}/L_\mathrm{bol}}$ compared to more active G stars.  Nevertheless, we can project the time series trends in their Figures 3 and 4 to recreate a qualitative history and fate of \hatpb~ under the assumptions of XUV photoevaporation.  Broadly, the planet would have started with an initial mass up to 50$\%$ greater than at present, with $R\sim\;$1.4 $R_\mathrm{Jup}$, for an initial density of $\sim$0.2 g/cm$^-3$.  It would lose mass at a rate of tens of Earth masses per Gyr for its 1 Gyr lifetime, expanding modestly until it reaches the critical $\sim$0.1 g/cm$^{-3}$ threshold, at which point the mass loss rate increases to its current value.  It will last only tens of Myr in its current state before losing almost all of its envelope and settling as a $5-15$ M$_\oplus$ core with a final radius of 0.2$-$0.3 R$_\mathrm{Jup}$.

@@

\begin{figure}
    \includegraphics[width=\linewidth]{figures/reinflation_MIST.png}
    \caption{Conceivable evolutionary scenarios for \hatpb.  The planet's equilibrium temperature increases as the stellar luminosity gradually increases on the main sequence.  The second panel from top shows the corresponding estimate for anomalous heating efficiency from \citet{2018AJ....155..214T}. The Ohmic dissipation model makes predictions for runaway inflation, depending on the incident stellar flux.  The heuristic scaling relation for the inflation timescale predicts vanishingly short inflationary lifetimes in this extreme regime but illustrates the countervailing effects of the top two panels.}
    \label{fig:OhmicInflate}
\end{figure}

>>>>

\begin{equation}
    \dot{M} = \frac{3}{4}\frac{\eta F_\mathrm{XUV}}{G K_t \rho_\mathrm{XUV}}\label{thorn23Mdot}.
\end{equation}
 where $K_t$ is the tidal gravity term, $\rho_\mathrm{XUV}$ is the planet density determined using the radius at which XUV radiation is deposited, and $\eta \sim 0.4$ comes from \citet{2022A&A...663A.122C}.  Assuming $\log{L_\mathrm{XUV}/L_\mathrm{bol}}=-4.2$, we would expect \hatpb~ to exhibit $\dot{M}\sim10^2$ M$_\oplus$/Gyr. At a current $M\sim95\;$M$_\oplus$ and an accelerating mass loss rate, its lifetime would measure in the $\sim$100 Myr range.  The \citet{2023ApJ...945L..36T} simulations focused on 0.75-1.25 $M_\odot$ host stars; the higher mass $\sim$1.6 $M_\odot$ \hatp would deliver greater tidal gravity for a given semi-major axis, and possibly lower $\log{L_\mathrm{XUV}/L_\mathrm{bol}}$ compared to more active G stars.  Nevertheless, we can project the time series trends in their Figures 3 and 4 to recreate a qualitative history and fate of \hatpb~ under the assumptions of XUV photoevaporation.  Broadly, the planet would have started with an initial mass up to 50$\%$ greater than at present, with $R\sim\;$1.4 $R_\mathrm{Jup}$, for an initial density of $\sim$0.2 g/cm$^-3$.  It would lose mass at a rate of tens of Earth masses per Gyr for its 1 Gyr lifetime, expanding modestly until it reaches the critical $\sim$0.1 g/cm$^{-3}$ threshold, at which point the mass loss rate increases to its current value.  It will last only tens of Myr in its current state before losing almost all of its envelope and settling as a $5-15$ M$_\oplus$ core with a final radius of 0.2$-$0.3 R$_\mathrm{Jup}$.

@@

\begin{figure}
    \includegraphics[width=\linewidth]{figures/reinflation_MIST2.png}
    \caption{Conceivable evolutionary scenarios for \hatpb.  The planet's equilibrium temperature increases as the stellar luminosity gradually increases on the main sequence.  The second panel from top shows the corresponding estimate for anomalous heating efficiency from \citet{2018AJ....155..214T}. The Ohmic dissipation model makes predictions for runaway inflation, depending on the incident stellar flux.  The heuristic scaling relation for the inflation timescale predicts vanishingly short inflationary lifetimes in this extreme regime but illustrates the countervailing effects of the top two panels.}
    \label{fig:OhmicInflate}
\end{figure}
